### 2018-03-21 paper/sexSelectionSpacing-ver1.tex
Better explanation of why changes within spell interesting

% [Why India?]
I apply the method to birth histories for Hindu women, using data from 
India's National Family and Health Surveys (NFHS), covering the period 
1972 to 2016. 
India is a particularly compelling case.
On the one hand, India has seen dramatic increases in the males-to-females ratio 
at birth over the last three decades as access to prenatal sex determination 
expanded 
\citep{das_gupta97,Sudha1999,Arnold2002,retherford03b,jha06}.%
\footnote{
India is not alone; both China and South Korea saw
significant changes in the sex ratio at birth over the same period 
\citep{Yi1993,park95}.
}
On the other hand, research suggests that son preference in 
India, when measured as ideally having more boys than girls, is decreasing 
over time and with higher education \citep{bhat03,pande07}.
% %
% \footnote{
% This measure of son preference is commonly used in the literature. 
% See, for example, \citet{clark00}, \citet{Jensen2009}, and \cite{Hu2015}.
% }
There have also been substantial changes in access and legality of
prenatal sex determination in India over the period covered.
Abortion has been legal in India since 1971 and still is, but the 
first reports of prenatal sex determination did
not appear until 1982--83 \citep{Sudha1999,bhat06,Grover2006}.
The number of clinics quickly increased, and knowledge about sex selection 
became widespread after a senior government official's wife aborted a 
fetus that turned out to be male \citep[p.\ 598]{Sudha1999}.
In 1994, the Central Government passed the Prenatal Diagnostic Techniques 
(PNDT) Act, making determining and communicating the sex of a fetus illegal.%
\footnote{
Details about the act are at \href{http://pndt.gov.in/}{http://pndt.gov.in/}.
The number of convictions has been low.
It took until January 2008 for the first state, Haryana, to reach five convictions.
Hence, private clinics apparently operate with little risk of legal action 
\citep{Sudha1999}.
Maharashtra was the first state to pass a similar law in 1988.
}
Hence, the data make it possible to show how the spacing between births and
sex ratios have changed with the introduction of prenatal sex determination,
and whether the ban affected the relationship between birth spacing and sex 
ratios.

@@

Finally, do parents change their decision on sex selection within a spell?
There are two reasons why this is interesting.
If parents change their mind the standard method of observing sex ratios 
will not capture the full extent of the use of sex selection.
Furthermore, it tells us something about parents' preferences for ideal
spacing between births and how strong their son preference is.

@@

Second, if the degree of son preference across women within the sample
but behavior is not affected by duration from the previous birth,
we should observe an increasing share of births come from 
parents with stronger son preference as duration from the
previous birth increases.
Whereas the early births are a combination of births to
people with lower son preference and those with stronger son
preference who conceived a son in the first or second pregnancy,
later births are then predominately from those with stronger
son preference who have not yet conceived a son.
Hence, in this case we should see an increase in the sex ratio 
as duration increases.

>>>>

% [Why India?]
I apply the method to birth histories for Hindu women, using data from 
India's National Family and Health Surveys (NFHS), covering the period 
1972 to 2016. 
India is a particularly compelling case.
On the one hand, India has seen dramatic increases in the males-to-females ratio 
at birth over the last three decades as access to prenatal sex determination 
expanded 
\citep{das_gupta97,Sudha1999,Arnold2002,retherford03b,jha06}.%
\footnote{
India is not alone; both China and South Korea saw
significant changes in the sex ratio at birth over the same period 
\citep{Yi1993,park95}.
}
On the other hand, research suggests that son preference in 
India, when measured as ideally having more boys than girls, is decreasing 
over time and with higher education \citep{bhat03,pande07}.
% %
% \footnote{
% This measure of son preference is commonly used in the literature. 
% See, for example, \citet{clark00}, \citet{Jensen2009}, and \cite{Hu2015}.
% }
There have also been substantial changes in access and legality of
prenatal sex determination in India over the period covered.
Abortion has been legal in India since 1971 and still is, but the 
first reports of prenatal sex determination did
not appear until 1982--83 \citep{Sudha1999,bhat06,Grover2006}.
The number of clinics quickly increased, and knowledge about sex selection 
became widespread after a senior government official's wife aborted a 
fetus that turned out to be male \citep[p.\ 598]{Sudha1999}.
In 1994, the Central Government passed the Prenatal Diagnostic Techniques 
(PNDT) Act, making determining and communicating the sex of a fetus illegal.%
\footnote{
Details about the act are at \href{http://pndt.gov.in/}{http://pndt.gov.in/}.
The number of convictions has been low.
It took until January 2008 for the first state, Haryana, to reach five convictions.
Hence, private clinics apparently operate with little risk of legal action 
\citep{Sudha1999}.
Furthermore, there is little evidence that bans like this significantly
affect sex ratios \citep{Das-Gupta2016}.
}
Hence, the data make it possible to show how the spacing between births and
sex ratios have changed with the introduction of prenatal sex determination,
and whether the ban affected the relationship between birth spacing and sex 
ratios.

@@

Finally, does the sex ratio change within a spell?
There are two reasons why this is interesting.
First, if the sex ratio changes within a spell simply using the observed sex
ratio at birth---as is standard practice in the literature---will not provide
correct estimates of the use of sex selection or the sex ratio when child
bearing is completed.
Second, the development in the sex ratio within a spell provides clues
to parents' preferences over the ideal spacing between births and the
distribution of son preference in the sample.

@@

Second, if the degree of son preference across women within the sample
but behavior is not affex2cted by duration from the previous birth,
we should observe an increasing share of births come from 
parents with stronger son preference as duration from the
previous birth increases.
Whereas the early births are a combination of births to
people with lower son preference and those with stronger son
preference who conceived a son in the first or second pregnancy,
later births are then predominately from those with stronger
son preference who have not yet conceived a son.
Hence, in this case we should see an increase in the sex ratio 
as duration increases.
